---
title: "Accessibility analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessibility analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 5)
```

```{r setup}
library(osmnxr)
```

This article walks through the kind of analysis `osmnxr` is built for: who can
reach a service, and how quickly. It underpins studies of urban mobility, green
mobility ("Fluxo Verde"), and access to schools and hospitals. We use the
offline grid so it runs anywhere; on a real city the steps are identical.

```{r}
g <- example_osm_graph(n = 10, spacing = 100)
g <- ox_add_edge_travel_times(g)
```

## Travel-time isochrones from a facility

Suppose a hospital sits near the centre of the area. The set of locations
reachable within a time budget is its **isochrone**. With `travel_time` as the
weight, cutoffs are in seconds:

```{r}
hospital <- ox_nearest_nodes(g, x = 450, y = 450)
iso <- ox_isochrone(g, hospital, cutoffs = c(60, 120), weight = "travel_time")
iso[c("cutoff", "n_nodes")]
```

```{r}
plot(g, col = "grey85")
plot(sf::st_geometry(iso), add = TRUE, border = NA,
     col = grDevices::adjustcolor(c("#0d3b66", "#2a9d8f"), 0.4))
```

The polygons are nested service areas: everything in the inner polygon is within
one minute of the hospital, the outer within two.

## Distance from many origins to many destinations

To measure how far a set of "schools" is from a set of "homes", use a distance
matrix. Here every cell is travel time in seconds:

```{r}
homes   <- ox_nearest_nodes(g, x = c(0, 900, 0),   y = c(0, 0, 900))
schools <- ox_nearest_nodes(g, x = c(450, 100),    y = c(450, 800))
m <- ox_distance_matrix(g, from = homes, to = schools, weight = "travel_time")
round(m)
```

The nearest school for each home is the row minimum:

```{r}
apply(m, 1, min)
```

## Putting it together on real data

With network access, the same pipeline runs on a city and its real facilities:

```{r, eval = FALSE}
g <- ox_graph_from_place("Olinda, Brazil", network_type = "drive") |>
  ox_simplify() |>
  ox_add_edge_travel_times()

hospitals <- ox_features_from_place("Olinda, Brazil", tags = list(amenity = "hospital"))
xy <- sf::st_coordinates(hospitals)
origins <- ox_nearest_nodes(g, xy[, 1], xy[, 2])

# 10- and 20-minute service areas across all hospitals
ox_isochrone(g, origins, cutoffs = c(600, 1200), weight = "travel_time")
```

That polygon set is the basis for equity analysis: overlay population and you
can quantify how many people fall outside a 20-minute reach — the core question
in accessibility and territorial planning.
