---
title: "Features and points of interest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Features and points of interest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(osmnxr)
```

Beyond street networks, `osmnxr` downloads arbitrary OpenStreetMap features —
amenities, buildings, transit stops, shops — as tidy `sf` points. Because these
calls contact the live Overpass API, the chunks below are not executed when the
vignette is built; run them interactively.

## Tag filters

Features are selected with **tags**, given as a named list. Each entry is either
`TRUE` (the key with any value) or a character vector of allowed values:

```{r, eval = FALSE}
# all schools in a place
ox_features_from_place("Olinda, Brazil", tags = list(amenity = "school"))

# schools and hospitals together
ox_features_from_place(
  "Recife, Brazil",
  tags = list(amenity = c("school", "hospital"))
)

# every building (key present, any value)
ox_features_from_place("Olinda, Brazil", tags = list(building = TRUE))
```

## From a bounding box

When you already know the extent, query a bounding box
(`c(xmin, ymin, xmax, ymax)` in longitude/latitude) directly:

```{r, eval = FALSE}
bbox <- c(-34.91, -8.07, -34.87, -8.04)
pois <- ox_features_from_bbox(bbox, tags = list(amenity = c("pharmacy", "clinic")))
pois
```

The result is an `sf` of points with `osm_type`, `osm_id` and one column per tag
encountered, so it composes directly with `dplyr` and `sf`:

```{r, eval = FALSE}
library(dplyr)
pois |>
  count(amenity, sort = TRUE)
```

## Combining features with a network

Features and networks share the same CRS (EPSG:4326), so you can snap POIs to
the network and analyse access. For example, find the nearest street node to
each hospital and compute travel-time isochrones from them — see the
[Routing and isochrones](routing-and-isochrones.html) and
[Accessibility](accessibility.html) articles.

```{r, eval = FALSE}
g <- ox_graph_from_place("Olinda, Brazil", network_type = "drive")
hospitals <- ox_features_from_place("Olinda, Brazil", tags = list(amenity = "hospital"))
xy <- sf::st_coordinates(hospitals)
nodes <- ox_nearest_nodes(g, xy[, 1], xy[, 2])
```
