---
title: "Getting started with osmnxr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with osmnxr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(osmnxr)
```

## What is osmnxr?

`osmnxr` is *OSMnx for R*: it downloads, models, analyzes and visualizes
street networks from [OpenStreetMap](https://www.openstreetmap.org/). The
public API is tidyverse-friendly and returns
[`sf`](https://r-spatial.github.io/sf/) objects; the heavy graph computation
(routing, metrics, simplification) runs in a bundled **Rust core**.

The central object is the `osm_graph`: a pair of `sf` tables (nodes and edges)
plus metadata.

## Working offline

Every part of the analysis API works without network access on a synthetic
grid, which is handy for learning and for reproducible examples:

```{r}
g <- example_osm_graph(n = 5)
g
```

```{r}
plot(g, nodes = TRUE)
```

## Network statistics

```{r}
ox_basic_stats(g)
```

## Routing

Find the nearest nodes to two coordinates, then compute the shortest path
(Dijkstra, in Rust):

```{r}
from <- ox_nearest_nodes(g, x = 0,   y = 0)
to   <- ox_nearest_nodes(g, x = 400, y = 400)
ox_shortest_path(g, from, to)
```

Single-source distances to every node:

```{r}
head(ox_distances(g, from))
```

## Street orientation

The Shannon entropy of edge bearings summarises how ordered a street grid is —
low for a gridiron, high for an organic network:

```{r}
ox_orientation_entropy(g)
```

## Downloading real data

With network access, build a graph straight from a place name (this contacts
the OpenStreetMap Overpass API, so it is not run here):

```{r, eval = FALSE}
g <- ox_graph_from_place("Olinda, Brazil", network_type = "drive")
plot(g)
ox_basic_stats(g)
```

See `?ox_graph_from_bbox`, `?ox_graph_from_point` and `?ox_geocode` for the
other download entry points, and `ox_settings()` to configure endpoints,
timeouts and caching.
