---
title: "Routing and isochrones"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Routing and isochrones}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 5)
```

```{r setup}
library(osmnxr)
```

This article uses the offline synthetic grid so it runs anywhere. With a real
network from `ox_graph_from_place()` the workflow is identical.

```{r}
g <- example_osm_graph(n = 8, spacing = 100)
```

## Shortest paths

Snap coordinates to graph nodes, then route between them. Routing runs in the
Rust core (Dijkstra).

```{r}
from <- ox_nearest_nodes(g, x = 0,   y = 0)
to   <- ox_nearest_nodes(g, x = 700, y = 700)
ox_shortest_path(g, from, to)
```

### Route alternatives

`ox_k_shortest_paths()` returns the *k* loopless shortest paths (Yen's
algorithm) as a tidy tibble with a list-column of node ids:

```{r}
ox_k_shortest_paths(g, from, to, k = 3)
```

## Travel time instead of distance

Attach speeds by road class and derive per-edge travel times, then route by
time by passing `weight = "travel_time"`.

```{r}
g <- ox_add_edge_travel_times(g)
head(g$edges[c("highway", "length", "speed_kph", "travel_time")])

ox_shortest_path(g, from, to, weight = "travel_time")
```

## Distance matrices

Many-to-many shortest-path costs in one call:

```{r}
hubs <- ox_nearest_nodes(g, x = c(0, 700, 0), y = c(0, 0, 700))
ox_distance_matrix(g, from = hubs, to = hubs)
```

## Isochrones (service areas)

An isochrone is the area reachable from one or more origins within a cost
budget. `ox_isochrone()` finds the reachable nodes (Rust Dijkstra) and hulls
them; with `travel_time` as the weight, cutoffs are in seconds.

```{r}
center <- ox_nearest_nodes(g, x = 350, y = 350)
iso <- ox_isochrone(g, center, cutoffs = c(150, 350), weight = "length")
iso[c("cutoff", "n_nodes")]
```

```{r}
plot(g, col = "grey80")
plot(sf::st_geometry(iso), add = TRUE,
     col = grDevices::adjustcolor(c("#0d3b66", "#2a9d8f"), 0.4), border = NA)
```

This is the core of accessibility analysis: with origins at schools or
hospitals and `travel_time` cutoffs, the polygons describe who can reach a
service within, say, 10 or 20 minutes — the basis for "Fluxo Verde" and
territorial-planning studies.
