---
title: "Street orientation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Street orientation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 5, fig.height = 5)
has_ggplot <- requireNamespace("ggplot2", quietly = TRUE)
```

```{r setup}
library(osmnxr)
```

Street-orientation analysis asks how ordered a city's grid is. `osmnxr`
computes the compass bearing of every edge and summarises the distribution with
Shannon entropy, all in the Rust core.

## Bearings

```{r}
g <- example_osm_graph(n = 6, spacing = 100)
b <- ox_bearings(g)
table(round(b))
```

A regular grid points in only a few directions, so its bearings cluster.

## Orientation entropy

Entropy (in nats) is low for an ordered gridiron and high for a disordered,
organic network. The perfectly aligned grid sits near the low end:

```{r}
ox_orientation_entropy(g)
```

For comparison, a network whose streets ran in every direction equally would
approach `log(num_bins)`:

```{r}
log(36)
```

## Rose plot

`ox_plot_orientation()` draws the classic polar histogram of bearings (requires
`ggplot2`):

```{r, eval = has_ggplot}
ox_plot_orientation(g, num_bins = 36)
```

With a real network this is where the character of a city shows: download one
with `ox_graph_from_place()`, simplify it, and plot:

```{r, eval = FALSE}
city <- ox_graph_from_place("Manhattan, New York, USA", network_type = "drive")
city <- ox_simplify(city)
ox_orientation_entropy(city) # low: a famous grid
ox_plot_orientation(city)
```
