---
title: "Urban metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Urban metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 5)
```

```{r setup}
library(osmnxr)
```

`osmnxr` summarises a street network with a handful of metrics, all computed in
the Rust core. This article uses the offline grid; the workflow is identical on
a network from `ox_graph_from_place()`.

```{r}
g <- example_osm_graph(n = 8, spacing = 100)
```

## Basic statistics

`ox_basic_stats()` reports node and edge counts, total and mean edge length,
mean out-degree, self-loops and average circuity:

```{r}
ox_basic_stats(g)
```

## Circuity

Circuity is the ratio of street length to straight-line distance between edge
endpoints. A value of `1` means perfectly straight streets; higher values mean
more winding ones. A grid is exactly straight:

```{r}
ox_circuity(g)
```

## Centrality

Which intersections carry the most through-traffic, and which are most central?
`ox_centrality()` returns betweenness (Brandes' algorithm) and closeness:

```{r}
ct <- ox_centrality(g, normalized = TRUE)
ct[order(-ct$betweenness), ][1:5, ]
```

Join the scores back onto the nodes to map them:

```{r}
nodes <- g$nodes
nodes$betweenness <- ct$betweenness[match(nodes$osmid, ct$osmid)]
plot(g, col = "grey85")
plot(nodes["betweenness"], pch = 19, add = TRUE)
```

The most between-central nodes sit in the core of the grid — exactly where
shortest paths concentrate. On a real network this highlights structurally
important junctions for mobility and resilience planning.
